Global Optimization of the Analogue Method by Means of Genetic Algorithms

### 2015-11-17 Genetic algorithms - Journal of Climate/template.tex
Removing redundant analyses.

The performance of the couples selection operator are relatively close (Figure \ref{fig:operator_selectcoupl_score}). Overall, we observe that the random pairing is the least good operator and the tournament selection with 3~candidates is slightly superior to others. Convergence analysis (not shown) presents similar trends to the scores. The influence on the number of assessments (not shown) reveals that the operators providing the worst scores are also those who need a higher number of evaluations. This analysis invites us to reject the rank and random pairing, and to consider the roulette wheel weighting or tournament selections, these being substantially better in terms of convergence and number of necessary evaluations.

Analysis of crossover operators (Figure \ref{fig:operator_crossover_score}) reveal some slightly superior options, some inappropriate, and many averages. Among the inappropriate operators, we find first the heuristic crossover, which is also more demanding in terms of evaluation number, and the linear crossover. Binary-like crossover (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence (not shown). The two-point intersection is also relatively close. Other operators can also be considered usable.

@@

After analysis of the most relevant mutation operators, we can raise the following advice:

@@

	\item \textit{Chromosome of adaptive search radius} (new): unlike the previous one, our new operator is very robust, as it requires no option and is auto-adapting. It may be sometimes a little bit slower for simple problems, but does not require parametrization, which is an important advantage. It is interesting to notice that our insertion of an extra chromosome representing the search radius gives better performance than other self-adaptive operators (such as, for example, the chromosome of adaptive mutation rate).

@@

@@

The size of the population has an effect on the performance of the optimization (Figure \ref{fig:option_taillepop_score}). A bigger population leads to better results, but also to significantly longer optimizations. Indeed, the required number of evaluation, and thus the required time, is approximately proportional to the population size. The optimal size seems to depend on the complexity of the AM to optimize: a more complex AM (ie. with more degrees of freedom) requires a bigger population size. A rule of thumb based on different experiences is provided in the next section.

@@

\subsection{Synthesis}

This comparison required tens of thousands of optimizations that were performed on a cluster of the University of Lausanne. The results, detailed in \citet{Horton2012a} are synthesized hereafter:

\begin{itemize}
	\item \textbf{Population size $N$}: we found the following ranges to be accurate in average for the population size (number of parameter sets of the AM):

	$50<N<100$ for a very simple implementation of the AM (1 level of analogy with 2 pressure levels),

	$N\approx200$ for a bit more complex AM (1 level of analogy with 4 pressure levels, or 2 level of analogy with less pressure levels),

	$N\approx500$ for significantly more complex AMs (2-3 levels of analogy with 4 pressure levels for the atmospheric circulation, and 2 to 4 levels for the moisture analogy),

	We didn't find any improvement with $N>500$, the results were even surprisingly of a lower quality. However, this cannot be generalized and depends on the AM to optimize, and supposedly on the characteristics of the processes generating the precipitations in a given region.

	\item \textbf{Natural selection}: this operator has no significant influence, and both tested implementations work fine, with a slightly better performance for the ratio-elitism.

	\item \textbf{Selection of couples}: 6 variants of the couples selection were assessed. The performance of these variants are relatively close, both in terms of performance score, convergence, and number of evaluations. The random pairing performed the most poorly, when the tournament selection with 3 candidates was slightly superior. The roulette wheel weighting is not far behind, but it is less effective in terms of convergence and number of evaluations. This operator has not a significant role in our application.

	\item \textbf{Chromosomes crossover}: we compared 21 different options of the crossover operators. This analysis revealed some slightly better options, some bad ones, and many averages. Among the bad operators, we find first the heuristic crossover, which is also more demanding in number of evaluations, as well as the linear crossover. Binary-like crossovers (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence. The two points crossover is relatively close. Other operators can be considered usable, yet may not be optimal. Once again, this operator is not the key of the GAs parametrization.

	\item \textbf{Mutation}: In our application, the mutation operator has a leading effect and should be chosen with care. We identified 3 more efficient variants: \textit{Non-uniform mutation} \citep{Michalewicz1996} that performs well but is difficult to parametrize, the \textit{Multi-scale mutation} (new) which needs only 1 parameter, and the \textit{Chromosome of adaptive search radius} (new), which is the most robust option as it has no parameter and always performed well.

	It may be wise to perform multiple optimizations and to consider these three operators in parallel in order to obtain results from algorithms that are either sometimes more efficient or more robust. It is interesting to note that the three best techniques incorporate a notion of search distance. It is likely that this notion is the key to these algorithms, for our application, and allows them to initially explore the parameter domain, and then to converge. The search radius in fact directly represents the notion of transition between exploration and exploitation, in our opinion more than a possible evolution of mutation rates.

	\item \textbf{Other options}: A ratio of 50\% for the mating pool seems to be a good choice.

\section{Recommended parametrization of GAs}

>>>>

The performance of the couples selection operator are relatively close (Figure \ref{fig:operator_selectcoupl_score}). Overall, we observe that the random pairing performed the most poorly, whereas the tournament selection with 3~candidates is slightly superior to others. The roulette wheel weighting is almost equivalent, but is a bit less effective in terms of convergence and number of evaluations (not shown). This operator has not a significant role in our application.

Analysis of crossover operators (Figure \ref{fig:operator_crossover_score}) reveal some slightly superior options, some inappropriate, and many averages. Among the inappropriate operators, we find first the heuristic crossover, which is also more demanding in terms of evaluation number, as well as the linear crossover. Binary-like crossover (especially with 2 points of intersection, whether $\beta$ is shared or not) are significantly better than the others, especially in terms of convergence (not shown). The two-point crossover is also relatively close. Other operators can also be considered usable.

@@

After analysis of the most relevant mutation operators, we can raise the following advice (detailed parametrization are provided in section \ref{sec:recommendations}):

@@

	\item \textit{Chromosome of adaptive search radius} (new): unlike the previous one, our new operator is very robust, as it requires no option and is auto-adapting. It may be sometimes a little bit slower for simple problems, but does not require parametrization, which is an important advantage. It is interesting to notice that our insertion of an extra chromosome representing the search radius gives better performance than other self-adaptive operators (such as, for example, the chromosome of adaptive mutation rate). If one had to choose a single option for the mutation operator, we would recommend the chromosome of adaptive search radius, as it was proven effective and needs no parameter.

@@

In our application, the mutation operator has a leading effect and should be chosen with care. It may be wise to perform multiple optimizations and to consider these three operators in parallel in order to obtain results from algorithms that are either sometimes more efficient or more robust. It is interesting to note that the three best techniques incorporate a notion of search distance. It is likely that this notion is the key to these algorithms, for our application, and allows them to initially explore the parameter domain, and then to converge. The search radius in fact directly represents the notion of transition between exploration and exploitation, in our opinion more than a possible evolution of mutation rates.

@@

The size of the population ($N$) has an effect on the performance of the optimization (Figure \ref{fig:option_taillepop_score}). A bigger population leads to better results, but also to significantly longer optimizations. Indeed, the required number of evaluation, and thus the required time, is approximately proportional to the population size. The optimal size seems to depend on the complexity of the AM to optimize: a more complex AM (ie. with more degrees of freedom) requires a bigger population size. A rule of thumb based on different experiences (not shown here) is provided hereafter:

\begin{itemize}
	\item $50<N\leqslant100$ for a very simple implementation of the AM (1 level of analogy with 2 pressure levels),
	\item $N\approx200$ for a bit more complex AM (1 level of analogy with 4 pressure levels, or 2 level of analogy with less pressure levels),
	\item $N\approx500$ for significantly more complex AMs (2-3 levels of analogy with 4 pressure levels for the atmospheric circulation, and 2 to 4 levels for the moisture analogy).
\end{itemize}

We didn't find any improvement with $N>500$, the results were even surprisingly of a lower quality. However, this relies on a limited number of case studies and cannot be generalized. It depends on the AM to optimize, and supposedly on the characteristics of the processes generating the precipitations in a given region.

@@

\section{Recommended parametrization of GAs}
\label{sec:recommendations}
